---
title: "Group assignment 1 - Measuring and testing 'quality of earnings' metrics using Compustat and XBRL data"
author: "Omar Alminawi, Akiel Camil, Kevin Liu, Ruben Morin, Maksimilian Stajer"
date: "11/03/2023"
output: html_document
---

<style type="text/css">
  body{
  font-size: 11pt;
}
  blockquote{
  font-size: 11pt;
  font-style: italic;
}
  h1{
  font-size: 20pt;
}
  h2{
  font-size: 16pt;
}
  h3{
  font-size: 14pt;
}
</style>

```{r setup, include=FALSE}
library(tidyr)
library(dplyr)
library(DT)
# Options for the R-notebook
knitr::opts_chunk$set(fig.width = 8, fig.height = 5, echo = T, warning=FALSE, message=FALSE, comment=FALSE, results = F)
knitr::opts_knit$set(progress = TRUE, verbose = TRUE)
```



```{r results=T}
setwd("C:/Users/Jiket/OneDrive/Desktop/UNI/Financial models/Assignment 1")
load(file = "dataset.rda")
head(dataset,100) %>% 
  datatable( rownames = FALSE, options = list(pageLength = 5, scrollX=T) )
```


```{r results=T}
dataset <- dataset %>% mutate(cyear = format(datadate, format="%Y")) 
dataset %>% group_by(cyear) %>% 
  count(., sort = F) %>% 
  datatable( rownames = FALSE, options = list(pageLength = 5, scrollX=T) )
dataset %>% group_by(FFindustry48) %>% 
  count(., sort = F) %>% 
  datatable( rownames = FALSE, options = list(pageLength = 5, scrollX=T) )
```



# Part 1
```{r}
library(dplyr)
library(lubridate)
```

First we check for missing values then we set the relevant N/A values to 0 and remove missing value where we cannot set them to 0, and remove all the firms whose financial years do not end on 31.12 and have missing years.
```{r}
# Identify columns with any missing values
columns_with_missing <- dataset %>% 
  summarize_all(~any(is.na(.))) %>%
  select_if(~any(.))

# Print the names of columns with missing values
print(names(columns_with_missing))


dataset$am[is.na(dataset$am)] <- 0
dataset$rectr[is.na(dataset$rectr)] <- 0
dataset$dp[is.na(dataset$dp)] <- 0
dataset$dlc[is.na(dataset$dlc)] <- 0
dataset$dltt[is.na(dataset$dltt)] <- 0

dataset$datadate <- as.Date(dataset$datadate, format = "%Y-%m-%d")

typeof(dataset$datadate)

dataset <- dataset %>% 
  filter(month(datadate) == 12 & day(datadate) == 31)

dataset$cyear <- as.numeric(dataset$cyear)

dataset <- dataset %>%
  filter(!is.na(xsga) & !is.na(oancf) & !is.na(rect) & !is.na(act) & !is.na(ppent))
dataset <- dataset %>%
  filter(rect != 0)

single_year_firms <- dataset %>%
  group_by(conm) %>%
  summarize(num_years = n()) %>%
  filter(num_years == 1) %>%
  pull(conm)

firms_with_gaps <- dataset %>%
  group_by(conm) %>%
  filter(!all(diff(cyear) == 1)) %>%
  pull(conm)

firms_to_remove <- unique(c(single_year_firms, firms_with_gaps))

data_final <- dataset %>%
  filter(!conm %in% firms_to_remove)
```
Most of the missing values that were removed came from xsga.

Next we will create a smaller data set with only the variables required to calculate the Beneish scores.

```{r}
beneish_data <- data_final %>%
  select(gvkey, conm, sale, cogs, xsga, dp, am, ib, oancf, rect, act, ppent, 
         at, dlc, dltt,cyear, datadate)

# Creating new columns as needed
# For Depreciation (excluding amortization)
beneish_data$dp_ex_am <- beneish_data$dp - beneish_data$am

# For Total Debt (short-term plus long-term)
beneish_data$td <- beneish_data$dlc + beneish_data$dltt

# Now remove the original dp, am, dlc, and dltt columns as they are no longer needed
beneish_data <- beneish_data %>%
  select(-dp, -am, -dlc, -dltt)

# View the first few rows of the new subset
head(beneish_data)
```

Check foe missing values in the new data set.

```{r}

# Check for missing values in each column
missing_values <- beneish_data %>% 
  summarize_all(~sum(is.na(.)))

# Print the results
print(missing_values)


```

Creating the lags of the relevant variables so we can calculate the indices needed to calculate the Beneish scores.
```{r}
beneish_data <- arrange(beneish_data, gvkey, conm, datadate)

# Calculate lags within gvkey, company name groups
beneish_data <- beneish_data %>% group_by(gvkey, conm) %>% 
                mutate(lsale = lag(sale, n = 1)) |> 
                mutate(lcogs = lag(cogs, n = 1)) |> 
                mutate(lxsga = lag(xsga, n = 1)) |> 
                mutate(ldp_ex_am = lag(dp_ex_am, n = 1)) |> 
                mutate(lib = lag(ib, n = 1)) |> 
                mutate(loancf = lag(oancf, n = 1)) |> 
                mutate(lrect = lag(rect, n = 1)) |> 
                mutate(lact = lag(act, n = 1)) |> 
                mutate(lppent = lag(ppent, n = 1)) |> 
                mutate(lat = lag(at, n = 1)) |> 
                mutate(ltd = lag(td, n = 1))

```

We can now calculate the indices needed for the Beneish scores, namely we calculate the Days’ Sales in Receivables Index, Gross Margin Index, Asset Quality Index, Sales Growth Index, Depreciation Index, Sales, General and Administrative expenses Index, Leverage Index, and the Accruals.

```{r}
attach(beneish_data)
beneish_data$dsri <- (rect/sale)/(lrect/lsale)
beneish_data$gmi <- ((1 - lcogs)/lsale)/((1-cogs)/sale)
beneish_data$aqi <- (1 - (ppent + act/at))/(1-(lppent + lact/lat))
beneish_data$sgi <- sale/lsale
beneish_data$depi <- (ldp_ex_am/(ldp_ex_am+ lppent))/(dp_ex_am/(dp_ex_am+ppent))
beneish_data$sgai <- (xsga/sale)/(lxsga/lsale)
beneish_data$levi <- (td/at)/(ltd/lat)
beneish_data$accruals <- (ib - oancf)/at
detach(beneish_data)
```

Before winsorizing we first have to remove the first firm- year for all companies as the 
indices cannot be calculated for the first year of a company's data.
The N/A values for levi can be set to 1, as the firm has no debt.
Because some companies have values of 0 for depreciation, 
we will set the N/A values for depi to the mean depi of each company.

```{r}

# Check for missing values in each column
missing_values <- beneish_data %>% 
  summarize_all(~sum(is.na(.)))

# Print the results
print(missing_values)

missing_values_per_column <- colSums(is.na(beneish_data))
missing_values_per_column

# Remove the first firm-year for each company as due to lags the inidices
# cannot be calculated for the first year of a company's data. We choose
# to filter by one of the lags to remove all the N/A values that appeared
# due to the lags.
beneish_data <- beneish_data %>%
  filter(!is.na(lsale))

missing_values_per_column <- colSums(is.na(beneish_data))
missing_values_per_column

# We can deal with N/A values in levi by setting to 1

beneish_data$levi[is.na(beneish_data$levi)] <- 1

missing_values_per_column <- colSums(is.na(beneish_data))
missing_values_per_column

checkdata <- beneish_data %>%
  filter(is.na(aqi))


beneish_data <- beneish_data %>%
  group_by(gvkey) %>%
  mutate(depi = ifelse(is.na(depi), mean(depi, na.rm = TRUE), depi))

# Putting the remaining values to 1, as the firm has no depreciation

beneish_data$depi[is.na(beneish_data$depi)] <- 1

# Setting the remaining AQI N/A value to Inf and letting the winsorize function
# deal with it. The firms have total assets equal to current assets therefore
# the AQI cannot be calculated for these firms.

beneish_data$aqi[is.na(beneish_data$aqi)] <- Inf
```



I removed the labels from the dataset but this can be changed after
```{r}
library(labelled)
for (var in names(beneish_data)) {
  var_label(beneish_data[[var]]) <- NULL
}


```


We install and load the package DescTools to winsorize the indices at the 5% and 95% levels,
```{r}
library(DescTools)

# List of variables to winsorize
variables_to_winsorize <- c('dsri', 'gmi', 'aqi', 'sgi', 'depi', 'sgai', 
                            'levi', 'accruals')
# Apply the Winsorize function to each variable
beneish_data[variables_to_winsorize] <- lapply(beneish_data[variables_to_winsorize], 
                                          function(x) Winsorize(x, 
                                                        probs = c(0.05, 0.95)))
```


Finally we can calculate the Beneish scores and add them to the dataset.

```{r}
attach(beneish_data)
beneish_data$bm <- -4.84 + 0.92 * dsri + 0.528 * gmi + 0.404 * aqi - 0.892 *
  sgi + 0.115 * depi - 0.172 * sgai + 4.679 * accruals - 0.327 * levi
detach(beneish_data)
```

